The impact of cosmic variance on simulating weak lensing surveys

### 2014-10-01 Writeup/paper_v3.tex
go over old conversations and remove / update

\rachel{Overall comment: the figures need a {\em lot} of work.  First,  they should be in the same order as they are referenced in the  text.  Second, they need to be sensibly referenced in the text; you  can't say ``as we see\dots'' without having a clear statement that  we should be looking at Fig.~4 (or whatever).  In all  cases, the font size is way too small, so you should enlarge it.  The rule of thumb for MNRAS and other journals is that the font size  in the figures should appear the same size as the font in the text  of the paper.  (This is true for axis labels, tick labels, and text  in the legend.)  You also need to write good captions, because some  people will not really read the paper, just the captions.  MNRAS
         does not use the subfigure environment, so you should remove it.  Just have the figures on top of or next to each other, and then in  the caption refer to ``Top panel'' and ``Bottom panel'' or whatever.} 
         \arun{Results have too many figures and they overlap with the bibliography. I think we should put some of the figures explicitly after the references?}
 \rachel{No, just move the figures a bit earlier in the text.  It's okay for them to show up slightly before you actually discuss them.  The references really should be last.}
\section{Introduction}
\label{S:intro}

@@

Previous studies have shown that any shape measurement method that uses second moments cannot be independent of the morphology and substructure of the galaxies involved. ({\bf rephrased from GREAT3}).
Hence, it is important to check any method that is proposed to measure the shape of the galaxies to check against a simulation that generates images as realistic as possible.  
With this in mind, the GREAT3 challenge~\citep{great3} was held with the aim to test and faciliate methods for measuring weak gravitational lensing by analyzing astronomical images.
Software packages like GalSim\footnote{https://github.com/GalSim-developers/GalSim} can generate galaxy images fairly accurately, given realistic input parametes.
The training dataset used in the challenge comes from the COSMOS survey, which we describe in sufficient detail in \S\ref{S:data}.
While providing a good quality (high $S/N$) images to be trained over, there could be some bias introduced due to the survey being a narrow one.
It is known \arun{cite refs.} that the shape and morphology of galaxies depend very much on their local environment. Hence, local overdensities or underdensities
observed in the COSMOS' field of view (FOV) may cause the properties of the galaxies to be associated with the redshift, thus unable to distinguish between redshift evolution and large scale structure.

@@

% \item \texttt{GOOD\_ZPHOT\_SOURCE=1}: This cut requires that there be a good photometric redshift, which typically is equivalent to requiring that the galaxy not be located within
%                                       the masked regions of the Subaru $BVIz$ imaging used for photometric redshifts, and that it have a successful match against an object in the Subaru imaging
%                                       \arun{copied from Mandelbaum
%                                         2011.}
 \item \texttt{GOOD\_ZPHOT\_SOURCE =1}: This cut requires that the photometric redshifts be reliable and good enough to draw conclusions about the population. More details on what `good' refers to can be found in \cite{SHERA}
\rachel{Need to rephrase, at least a little.  I recommend being
  briefer and saying to see that paper for more details.} \arun{Done.}
\end{enumerate}
%\rachel{The cuts you just gave are really basic or fundamental ones.  Those should go before any discussion of more complicated derived quantities like redshifts, stellar mass, Sersic fits, etc.}

Precise shape measurements, when compared to ground-based surveys, can be made since the full width half-maximum (FWHM) of the point-spread function (PSF)
is $0.12''$. High resolution images taken through the wide F814W filter (broad \emph{I}), after correcting for the PSF, have led to collection of postage stamp images\footnote{http://irsa.ipac.caltech.edu/data/COSMOS/images/galaxy-postage-stamps/} for the GREAT3 challenge~\citep{great3}, using the procedure described in \cite{SHERA}.
Parametric models to most of these galaxies including \sersicn profile fits, 
2 component bulge+disk fits, axis ratios etc were done for GREAT3 challenge~\citep{great3} using the method described in \cite{Claire_Fits}. This is described briefly in Sec.~\ref{sub:axisratio}.In addition to the ACS/WFC (F814W) imaging, the COSMOS field has also been imaged by Subaru Suprime-Cam ($B_j, V_j, g^+,r^+,i^+,z^+,$ NB816), the
Canada-French Hawaii Telescope (CFHT;$u^*,i^*$) and the KPNO/CTIO ($K_s$) .

@@

%\rachel{I thought we were only going up to $z=0.85$ but here you say $1$?}
%For our sample of galaxies, volume-limited upto $z=1$ by imposing an $M_I$ cut, we fit parametric models to the histogram of photometric redshifts in order to assign values of overdensity. The figures show chi-squared and gamma functions fitted to the histogram~\citep{Redshift_modelling}.
For our (flux-limited) sample of galaxies, up to $z=1.0$, we fit parametric models to the histogram of photometric redshift in order to assign values of overdensity. Figures show Chi-squared distribution given by
\begin{equation}
 p_1(z) \sim z^{a-1}\exp({-bz})
\end{equation}
and the function given by
\begin{equation}
 p_2(z) \sim z^2\exp{\left(-\left(\frac{z}{z_0}\right)^{1.5}\right)}
\end{equation}
that was first presented by~\cite{Redshift_modelling} as fits to the
redshift histograms. The normalization constants depend not only on the parameters but also on the lower and the upper limit of the redshifts considered. 
\rachel{(To avoid ambiguity, you should use avoid   using the same symbol in both equations.  For example you could turn   the $b$'s into $b_1$ and $b_2$.  Also, the normalization factors out  front do not matter - they are not something that you fit for.  I would say $p(z)\propto$\dots to avoid   the need for a normalization for which you are not actually fitting.)} 
\arun{Done.}
Here $a,b$ and $z_0$ are free parameters that are to be determined.
%\rachel{You cannot expect people to look these up.  Give them equations in which you define $p(z)$ for each case, with labels like $p_1(z)$ and $p_2(z)$.}
For low $z$, the volume is too small to rely the overdensity values from our model fits and hence the fit is made for $z \ge 0.3$. Our binning width, $\Delta z$ must be small enough so as to be able to identify localized overdensities/underdensities but large enough to not let our conclusions be affected by the errors in photometric redshift. 
We choose our bins to be $0.05$ wide starting from $z=0.3$. 
%\rachel{Were you going to use other fitting functions as a test of the method?  If so, you should introduce them here.}

@@

Overdensity (as predicted by a given model) in a redshift bin is defined as the ratio of the
difference between the value given by the histogram and the value
predicted by the model to the latter:
\begin{equation}
\delta_{g,\text{1D}}=\frac{(N-N_{\text{mod}})}{N_{\text{mod}}}.  
\end{equation}
%(\rachel{This $\delta$ definition is important, so it should be in an equation environment.})
(\rachel{Call it $\delta_{g,\text{1D}}$, not $\delta$, which implies matter
  overdensity.  Better yet would be $\delta_{g,1D}$, to clearly
  indicate this is based on a 1-dimensional slice through redshift
  space rather than isolating 3D overdensities.})
Note that $\delta_{g,\text{1D}}$ is model dependent. If a model and the distribution agree, then $\delta_{g,\text{1D}}=0$ for that model at all the redshifts.

@@

COSMOS is a flux-limited survey and is therefore affected by Malmquist bias, i.e. at higher redshifts, the brighter galaxies are preferentially observed.
Our analysis involves comparing galaxies in different redshift slices to find significant differences in morphology, if any, so with a flux-limited sample, we would be comparing only the bright galaxies at high redshifts with
bright and faint galaxies at low redshifts. For a fair comparison, we must restrict ourselves to bright galaxies at all redshifts and this is acheived by volume-limiting the sample.
%\rachel{Explain why you are doing this.  Before saying {\em how} you
%  generate a volume-limited sample, explain {\em why} you are
%  generating one.}
We generate, using the method below, a volume-limited sample that is complete upto $z=1$ by applying a cut on luminosity such that only galaxies instrinsically brighter than a certain threshold is considered. This threshold is set on $K$-corrected $F814W$-band magnitudes, denoted as $M_\text{F814W}$.
\rachel{Is it $i$ band, or $I$ band?  Check which one and then be
  self-consistent throughout the text.}\arun{I think they call it
  $I_{814}$, so $I$.}
\rachel{But you are still being ambiguous in sometimes calling it $I$
  and other times F814W.  Pick one.}
Since the parent sample contains fainter galaxies, upto $m_\text{F814W}=25.2$, we compare the distribution of the $m_\text{F814W}=23.5$ sample with the samples containing fainter galaxies for high redshift bins, to see where the sample is no longer complete.
% \begin{figure}
%  \centering
%  a) \includegraphics[width=\columnwidth]{volume_limiting_pdf}
%  b) \includegraphics[width=\columnwidth]{volume_limiting_cdf}
%  \label{fig:volume_limiting_dist}
%  \caption{ ... }
% \end{figure}

@@

\rachel{In my opinion, given that the main results are from Claire's
  fits, you should describe those {\em first} and then put the
  re-Gaussianization description second.  Also, state this plan up
  front, i.e., something like ``To describe galaxy morphology and
  shape, we will use fits to Sersic profiles. (\dots description here
  \dots)  However, as a secondary check on the results, we will use \dots''}

@@

\section{Results}
\label{S:results}
Having identified the overdense and underdense regions in a volume-limited sample (\S\ref{sub:overdensities}, \S\ref{sub:volumelimiting}), we will now see in this section if the morphological
parameters of the galaxies, listed in \S\ref{sub:axisratio}, have any dependence on the environment that they reside in. We have 3 different ways of volume-limiting our sample
\begin{enumerate}
 \item no redshift evolution of luminosity cut $(\s1)$,
 \item use $B$-band evolution rate for $I$-band luminosities $(\s2)$,
 \item impose stellar mass cuts instead of luminosity $(\s3)$
\end{enumerate}
and we will present our results in each of the 3 cases. 
\rachel{Say how this connects to the 3 samples S1, S2, and S3 that you defined earlier.} \arun{Done.}

\rachel{I also recommend splitting this section up into subsections so
  it's not just one big block.  For example, the first can be about
  axis ratios (with a subsubsection about how the results depend on
  the way the volume limiting was carried out), the second can be
  about morphology from Sersic index and bulge-to-total ratios.}

@@

We first compare the distribution of the axis ratio in \emph{all}
overdense bins and \emph{all} underdense bins in Fig.~\ref{fig:axisratio_all}.
\rachel{How was the volume-limiting carried out for this?  You should say the answer to this question  for all figures.} 
\arun{The caption mentions it.}
Unless otherwise mentioned, the axis ratios refer to the values obtained using the method of~\cite{Claire_Fits}.
The cumulative distinction functions are also plotted alongside in order to be able to visualize the `distance' between the distributions.
Referring to the first two rows in Table~\ref{table:pvalues_all}, we see that the p-values from both the KS and AD tests are well below 0.05
(\rachel{give the actual numbers!} 
\arun{It's in the table.} 
\rachel{The text does not seem to mention the table at all.  If thereaders are to get the numbers from it, then you need to refer to it here and any other place where it is relevant.}), 
\arun{Done.}
so we reject the `null hypothesis' that the overdense and underdense regions have same axis ratio distributions at 95\% significance level.

@@

Such a trend is also observed in axis ratios based on second moments.\arun{I've commented on the agreement/disagreement of the distributions themselves in Table~\ref{table:pvalues_momentbased_all}. No plots included.} 
If $Q_{ij}$ represents the generic matrix element of the matrix of second moments, then define a complex number, sometimes known as the \emph{distortion}, as
\begin{equation}
 e = e_1 + ie_2 = \frac{Q_{11}-Q_{22}+2iQ_{12}}{Q_{11}+Q_{22}}.
 \label{eq:distortion}
\end{equation}
\arun{Is this the correct denominator?}
Then, one definition of ellipticity is the magnitude of this complex number which is $\sqrt{e_1^2+e_2^2}$.

>>>>

\section{Introduction}
\label{S:intro}

@@

Previous studies have shown that any shape measurement method that uses second moments cannot be independent of the morphology and substructure of the galaxies involved. ({\bf rephrased from GREAT3}).
Hence, it is important to check any method that is proposed to measure the shape of the galaxies to check against a simulation that generates images as realistic as possible.  
With this in mind, the GREAT3 challenge~\citep{great3} was held with the aim to test and faciliate methods for measuring weak gravitational lensing by analyzing astronomical images.
Software packages like GalSim\footnote{\url{https://github.com/GalSim-developers/GalSim}} can generate galaxy images fairly accurately, given realistic input parametes.
The training dataset used in the challenge comes from the COSMOS survey, which we describe in sufficient detail in \S\ref{S:data}.
While providing a good quality (high $S/N$) images to be trained over, there could be some bias introduced due to the survey being a narrow one.
It is known \arun{cite refs.} that the shape and morphology of galaxies depend very much on their local environment. Hence, local overdensities or underdensities
observed in the COSMOS' field of view (FOV) may cause the properties of the galaxies to be associated with the redshift, thus unable to distinguish between redshift evolution and large scale structure.

@@

% \item \texttt{GOOD\_ZPHOT\_SOURCE=1}: This cut requires that there be a good photometric redshift, which typically is equivalent to requiring that the galaxy not be located within
%                                       the masked regions of the Subaru $BVIz$ imaging used for photometric redshifts, and that it have a successful match against an object in the Subaru imaging
%                                       \arun{copied from Mandelbaum
%                                         2011.}
 \item \texttt{GOOD\_ZPHOT\_SOURCE =1}: This cut requires that the photometric redshifts be reliable and good enough to draw conclusions about the population. More details on what `good' refers to can be found in \cite{SHERA}
\end{enumerate}
%\rachel{The cuts you just gave are really basic or fundamental ones.  Those should go before any discussion of more complicated derived quantities like redshifts, stellar mass, Sersic fits, etc.}

Precise shape measurements, when compared to ground-based surveys, can be made since the full width half-maximum (FWHM) of the point-spread function (PSF)
is $0.12''$. High resolution images taken through the wide F814W filter (broad \emph{I}), after correcting for the PSF, have led to collection of postage stamp images\footnote{\url{http://irsa.ipac.caltech.edu/data/COSMOS/images/galaxy-postage-stamps/}} for the GREAT3 challenge~\citep{great3}, using the procedure described in \cite{SHERA}.
Parametric models to most of these galaxies including \sersicn profile fits, 
2 component bulge+disk fits, axis ratios etc were done for GREAT3 challenge~\citep{great3} using the method described in \cite{Claire_Fits}. This is described briefly in Sec.~\ref{sub:axisratio}.In addition to the ACS/WFC (F814W) imaging, the COSMOS field has also been imaged by Subaru Suprime-Cam ($B_j, V_j, g^+,r^+,i^+,z^+,$ NB816), the
Canada-French Hawaii Telescope (CFHT;$u^*,i^*$) and the KPNO/CTIO ($K_s$) .

@@

%\rachel{I thought we were only going up to $z=0.85$ but here you say $1$?}
%For our sample of galaxies, volume-limited upto $z=1$ by imposing an $M_I$ cut, we fit parametric models to the histogram of photometric redshifts in order to assign values of overdensity. The figures show chi-squared and gamma functions fitted to the histogram~\citep{Redshift_modelling}.
For our (flux-limited) sample of galaxies, up to $z=1.0$, we fit parametric models to the histogram of photometric redshift in order to assign values of overdensity. Figures show Chi-squared distribution given by
\begin{equation}
 p_1(z) \sim z^{a-1}\exp({-bz})
\end{equation}
and the function given by
\begin{equation}
 p_2(z) \sim z^2\exp{\left(-\left(\frac{z}{z_0}\right)^{1.5}\right)}
\end{equation}
that was first presented by~\cite{Redshift_modelling} as fits to the
redshift histograms. The normalization constants depend not only on the parameters but also on the lower and the upper limit of the redshifts considered. 
Here $a,b$ and $z_0$ are free parameters that are to be determined.
%\rachel{You cannot expect people to look these up.  Give them equations in which you define $p(z)$ for each case, with labels like $p_1(z)$ and $p_2(z)$.}
For low $z$, the volume is too small to rely the overdensity values from our model fits and hence the fit is made for $z \ge 0.3$. Our binning width, $\Delta z$ must be small enough so as to be able to identify localized overdensities/underdensities but large enough to not let our conclusions be affected by the errors in photometric redshift. 
We choose our bins to be $0.05$ wide starting from $z=0.3$. 
%\rachel{Were you going to use other fitting functions as a test of the method?  If so, you should introduce them here.}

@@

Overdensity (as predicted by a given model) in a redshift bin is defined as the ratio of the
difference between the value given by the histogram and the value
predicted by the model to the latter:
\begin{equation}
\delta_{g,\text{1D}}=\frac{(N-N_{\text{mod}})}{N_{\text{mod}}}.  
\end{equation}
%(\rachel{This $\delta$ definition is important, so it should be in an equation environment.})
Note that $\delta_{g,\text{1D}}$ is model dependent. If a model and the distribution agree, then $\delta_{g,\text{1D}}=0$ for that model at all the redshifts.

@@

COSMOS is a flux-limited survey and is therefore affected by Malmquist bias, i.e. at higher redshifts, the brighter galaxies are preferentially observed.
Our analysis involves comparing galaxies in different redshift slices to find significant differences in morphology, if any, so with a flux-limited sample, we would be comparing only the bright galaxies at high redshifts with
bright and faint galaxies at low redshifts. For a fair comparison, we must restrict ourselves to bright galaxies at all redshifts and this is acheived by volume-limiting the sample.
%\rachel{Explain why you are doing this.  Before saying {\em how} you
%  generate a volume-limited sample, explain {\em why} you are
%  generating one.}
We generate, using the method below, a volume-limited sample that is complete upto $z=1$ by applying a cut on luminosity such that only galaxies instrinsically brighter than a certain threshold is considered. This threshold is set on $K$-corrected $F814W$-band magnitudes, denoted as $M_\text{F814W}$.
\rachel{Is it $i$ band, or $I$ band?  Check which one and then be
  self-consistent throughout the text.}\arun{I think they call it
  $I_{814}$, so $I$.}
\rachel{But you are still being ambiguous in sometimes calling it $I$
  and other times F814W.  Pick one.}
\rachel{Actually, is this the catalog entry called ``MI''?  If so, it really is a different band - it's the k-corrected $I$-band data from Suburu or CFHT, so it's appropriate to call it $M_I$ and explain briefly that it's from the Subaru data (but $m_{F814W}$, since MAG\_AUTO is from F814W).}
Since the parent sample contains fainter galaxies, upto $m_\text{F814W}=25.2$, we compare the distribution of the $m_\text{F814W}=23.5$ sample with the samples containing fainter galaxies for high redshift bins, to see where the sample is no longer complete.
% \begin{figure}
%  \centering
%  a) \includegraphics[width=\columnwidth]{volume_limiting_pdf}
%  b) \includegraphics[width=\columnwidth]{volume_limiting_cdf}
%  \label{fig:volume_limiting_dist}
%  \caption{ ... }
% \end{figure}

@@

@@

\section{Results}
\label{S:results}
Having identified the overdense and underdense regions in a volume-limited sample (\S\ref{sub:overdensities}, \S\ref{sub:volumelimiting}), we will now see in this section if the morphological
parameters of the galaxies, listed in \S\ref{sub:axisratio}, have any dependence on the environment that they reside in. We have 3 different ways of volume-limiting our sample
\begin{enumerate}
 \item no redshift evolution of luminosity cut $(\s1)$,
 \item use $B$-band evolution rate for $I$-band luminosities $(\s2)$,
 \item impose stellar mass cuts instead of luminosity $(\s3)$
\end{enumerate}
and we will present our results in each of the 3 cases.

@@

We first compare the distribution of the axis ratio in \emph{all}
overdense bins and \emph{all} underdense bins in Fig.~\ref{fig:axisratio_all}.
\rachel{How was the volume-limiting carried out for this?  You should say the answer to this question  for all figures.} 
\arun{The caption mentions it.}
\rachel{I'm afraid that's not enough.  Some people are going to read linearly and you want the info in the main text, too.}
Unless otherwise mentioned, the axis ratios refer to the values obtained using the method of~\cite{Claire_Fits}.
The cumulative distinction functions are also plotted alongside in order to be able to visualize the `distance' between the distributions.
Referring to the first two rows in Table~\ref{table:pvalues_all}, we see that the p-values from both the KS and AD tests are well below 0.05, 
so we reject the `null hypothesis' that the overdense and underdense regions have same axis ratio distributions at 95\% significance level.

@@

Such a trend is also observed in axis ratios based on second moments. 
If $Q_{ij}$ represents the generic matrix element of the matrix of second moments, then define a complex number, sometimes known as the \emph{distortion}, as
\begin{equation}
 e = e_1 + ie_2 = \frac{Q_{11}-Q_{22}+2iQ_{12}}{Q_{11}+Q_{22}}.
 \label{eq:distortion}
\end{equation}
\arun{Is this the correct denominator?} \rachel{Yes.  There's another ellipticity definition with a different denominator, but we use distortion which is defined as in this equation.  (Once you read this response feel free to delete this conversation.)}
Then, one definition of ellipticity is the magnitude of this complex number which is $\sqrt{e_1^2+e_2^2}$.
